SU($n$) ferromagnets near the Heisenberg point

### 2020-06-19 collective_sun/sun_perturbation.tex
minor wording changes

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    \draw[thickest] (0,0) -- ++(10em,0);
    \foreach \hh in {0.15em,0.05em,-0.05em,-0.15em} {
      \draw (0,3em+\hh) -- ++(10em,0);
    }
    \draw[->] (7em,.25em) arc (220:0:.4em);
    \draw[->] (7em,.25em) to [bend left]
    node[left] {$\O\p{\m v,\sigma_\z}$} (7em,2.7em);
    \node () at (5em,4.7em) {$\vdots$};
    \foreach \hh in {0.25em,0.15em,0.05em,-0.05em,-0.15em,-0.25em} {
      \draw (0,6em+\hh) -- ++(10em,0);
    }
    \draw[->] (-.4em,0) arc (220:0:.4em);
    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,3em);
    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,4.3em);
    \draw[->] (-.4em,0) to [bend left, in=130]
    node[left] {$\O\p{\m w,\sigma_\z^{\otimes r}}$} (-.2em,6em);
    \node[right] () at (10em,0) {$\M_0,S=\frac{N}{2}$};
    \node[right] () at (10em,3em) {$\M_1,S=\frac{N}{2}-1$};
    \node[right] () at (10em,6em) {$\M_r,S=\frac{N}{2}-r$};
  \end{tikzpicture}
  \caption{Schematic of relationships between $q$-magnon manifolds
    $\M_q$ with total spin $S=N/2-q$ in an SU(2) spin system.
    Single-body operators $\O\p{\m v,\sigma_\z}$ generally couple the
    ground-state manifold $\M_0$ to the single-magnon manifold $\M_1$
    as well as $\M_0$ itself, while $r$-body operators
    $\O\p{\m w,\sigma_\z^{\otimes r}}$ couple $\M_0$ to all $q$-magnon
    manifolds $\M_q$ with $q\le r$.}
  \label{fig:manifold_sketch}
\end{figure}

In the case of an SU($n$) spin system, we can similarly start with the
ground-state PS manifold
$\M_0=\oper{span}\set{\ket{m}:m\in\A_n\p{N}}$, and generate excited
states by the action of multi-body operators, $\O\p{\m w,X}\ket{m}$.
The choice of tensors $\m w$ for generating states in the $r$-magnon
manifold $\M_r$ is the same as before, but we now have the additional
complication of having to choose $r$-spin operators, $X$.  As with the
choice of $\sigma_\z^{\otimes r}$ for SU(2), it suffices to choose
tensor products of {\it diagonal} single-spin operators (that is,
diagonal in a fixed single-particle basis), but there will now be
$\sim\p{n-1}^r$ choices: one of $n-1$ diagonal operators (that are
linearly independent of the identity) for each of $r$ tensor factors.
We discuss the exact choice of operators in more detail in Appendix
\ref{sec:operator_content}.  To summarize, we define $n$ single-spin
operators $Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$, for
which $Z_\nu=\pm\sigma_\z$ in the case of SU(2), and we choose a list
of non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with
$\sum_\mu z_\mu=r$ to define the $r$-spin operator
\begin{align}
  X_z \equiv \bigotimes_{\nu=1}^n Z_\nu^{\otimes z_\nu}
  = \underbrace{Z_1 \otimes \cdots \otimes Z_1}_{z_1~\t{times}}
  \otimes
  \underbrace{Z_2 \otimes \cdots \otimes Z_2}_{z_2~\t{times}}
  \otimes \cdots \otimes
  \underbrace{Z_n \otimes \cdots \otimes Z_n}_{z_n~\t{times}}.
\end{align}
Our starting basis for $\M_r$ then consists of states
$\ket{m,\m w,X_z}\equiv\O\p{\m w,X_z}\ket{m}$, where we enforce that
$z_\nu\le m_\nu$ for all $\nu$ (where $m_\nu$ is the occupation number
of the single-spin state $\nu$), and that $z$ is zero in an entry
where $m$ is maximal: $z_\nu=0$ for
$\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  This second restriction is
equivalent to removing an operator from $\set{Z_\nu:\nu\in\ZZ_n}$,
making the remaining set linearly independent of the identity
operator.  The basis $\set{\ket{m,\m w,X_z}}$ for $\M_r$ is generally
complete, but not orthonormal, so as a final step we use the states
$\ket{m,\m w,X_z}$ to (numerically) construct an orthonormal basis
$\B_r$ for $\M_r$ (see Appendix \ref{sec:operator_content}).

@@

Next, we address the fact that the basis $\set{\ket{m,\m w,z}}$ for
$\M_r$ is overcomplete.  This fact is a simple consequence of our
choice of single-spin operators for the construction of
$\ket{m,\m w,z}$, namely $\set{Z_\nu:\nu\in\ZZ_n}$.  In essence, our
problem is the fact that $Z_\nu=\p{n-2}-\sum_{\mu\ne\nu}Z_\mu$, which
implies that any state $\ket{m,\m w,z}\in\M_r$ can be written as a
linear combination of states in which $z_\nu\to z_\nu-1$, and possibly
$z_\mu\to z_\mu+1$ for some $\mu$.  We can eliminate this redundancy
by removing an operator from the $\set{Z_\nu}$, or equivalently
enforcing that $z_\nu=0$ for some $\nu$.  As a matter of choice, we
therefore consider only states $\ket{m,\m w,z}$ for which $z$ is zero
at the first entry where $m$ is maximal: $z_\nu=0$ for
$\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  Given our previous restriction
that $z_\nu\le m_\nu$ for all $\nu$, enforcing that $z$ is zero at an
entry where $m$ is maximal maximizes the number of candidate values of
$z$ that we eliminate from consideration.

@@

To simplify $\Q\p{\m w,\m X}$, we first collect terms to express
\eqref{eq:sym_prod_start} as a sum of products:
\begin{align}
  \Q\p{\m w,\m X} = \sum_{k\in\D_N\p{\m M}} \prod_{\alpha\in\ZZ_p}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
  \label{eq:sym_prod_sum}
\end{align}
where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the dimensions
$M_\alpha$; and
$\D_N\p{\m M}\equiv\bigotimes_{M_\alpha\in\m M}\D_N\p{M_\alpha}$, such
that each $k\in\D_N\p{\m M}$ has the decomposition
$k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\D_N\p{M_\alpha}$.  We
then classify the terms in \eqref{eq:sym_prod_sum} by the numbers
$g_S$ of indices shared by all tensors $w_\alpha$ with
$\alpha\in S\subset\ZZ_p$.  For example, a term of the form
$w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
$a,b,c,d,e$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,3}} = 1,
  &&
  g_{\set{2,3}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be associated with the Venn diagram
\begin{align}
  \diagram{example_123},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
circles $\alpha\in S$.  We denote the set of all nonempty subsets of
$\ZZ_p$ by $\PP^*\p{\ZZ_p}$, and denote a choice of $g_S$ for all
$S\in\PP^*\p{\ZZ_p}$ by $g$.  With representations such as
\eqref{eq:venn_diagram} in mind, we will loosely refer to $g$ as
simply a ``diagram''.  We define
$\abs{g}\equiv\sum_{S\in\PP^*\p{\ZZ_p}}g_S$ to be the number of
indices considered in $g$ (or equivalently the number of dots in the
diagram $g$), and we denote the set of all diagrams for
$p=\dim\p{\m M}$ tensors with dimensions $\m M$ by $\G\p{\m M}$.  For
consistency, all diagrams $g\in\G\p{\m M}$ must associate exactly
$M_\alpha$ indices with the dimension-$M_\alpha$ tensor $w_\alpha$,
i.e.
\begin{align}
  \sum_{S\in\PP^*\p{\ZZ_p}\,:\,\alpha\in S} g_S
  = \sum_{R\in\PP^*\p{\ZZ_p\setminus\set{\alpha}}} g_{\set{\alpha}\cup R}
  = M_\alpha
\end{align}
for all $\alpha\in\ZZ_p$.  A classification of the terms in
\eqref{eq:sym_prod_start} by diagrams $g\in\G\p{\m M}$ allows us to
expand
\begin{align}
  \Q\p{\m w,\m X} \EQPS \sum_{g\in\G\p{\m M}} \Q_g\p{\m w,\m X},
  \label{eq:sym_prod_diagrams}
\end{align}
where $\EQPS$ denotes equality under a restriction to $\M_0$, and
$\Q_g\p{\m w,\m X}$ is the sum of all terms in
\eqref{eq:sym_prod_start} that are associated with the diagram $g$.

>>>>

% focus on energy -- how different from non-interacting magnons?
% plot dispersion relation
% lax formulation for single-particle terms

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    \draw[thickest] (0,0) -- ++(10em,0);
    \foreach \hh in {0.15em,0.05em,-0.05em,-0.15em} {
      \draw (0,3em+\hh) -- ++(10em,0);
    }
    \draw[->] (7em,.25em) arc (220:0:.4em);
    \draw[->] (7em,.25em) to [bend left]
    node[left] {$\O\p{\m v,\sigma_\z}$} (7em,2.7em);
    \node () at (5em,4.7em) {$\vdots$};
    \foreach \hh in {0.25em,0.15em,0.05em,-0.05em,-0.15em,-0.25em} {
      \draw (0,6em+\hh) -- ++(10em,0);
    }
    \draw[->] (-.4em,0) arc (220:0:.4em);
    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,3em);
    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,4.3em);
    \draw[->] (-.4em,0) to [bend left, in=130]
    node[left] {$\O\p{\m w,\sigma_\z^{\otimes r}}$} (-.2em,6em);
    \node[right] () at (10em,0) {$\M_0,S=\frac{N}{2}$};
    \node[right] () at (10em,3em) {$\M_1,S=\frac{N}{2}-1$};
    \node[right] () at (10em,6em) {$\M_r,S=\frac{N}{2}-r$};
  \end{tikzpicture}
  \caption{Schematic of relationships between $q$-magnon manifolds
    $\M_q$ with total spin $S=N/2-q$ in an SU(2) spin system.
    Single-body operators $\O\p{\m v,\sigma_\z}$ generally couple the
    ground-state manifold $\M_0$ to the single-magnon manifold $\M_1$
    as well as $\M_0$ itself, while $r$-body operators
    $\O\p{\m w,\sigma_\z^{\otimes r}}$ generally couple $\M_0$ to all
    $q$-magnon manifolds $\M_q$ with $q\le r$.}
  \label{fig:manifold_sketch}
\end{figure}

In the case of an SU($n$) spin system, we can similarly start with the
ground-state PS manifold
$\M_0=\oper{span}\set{\ket{m}:m\in\A_n\p{N}}$, and generate excited
states by the action of multi-body operators, $\O\p{\m w,X}\ket{m}$.
The choice of tensors $\m w$ for generating states in the $r$-magnon
manifold $\M_r$ is the same as before, but we now have the additional
complication of having to choose $r$-spin operators, $X$.  As with the
choice of $\sigma_\z^{\otimes r}$ for SU(2), it suffices to choose
tensor products of {\it diagonal} single-spin operators (that is,
diagonal in a fixed single-particle basis), but there will now be
$\sim\p{n-1}^r$ choices: one of $n-1$ diagonal operators (that are
linearly independent of the identity) for each of $r$ tensor factors.
We discuss the exact choice of operators in more detail in Appendix
\ref{sec:operator_content}.  To summarize, we define $n$ single-spin
operators $Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$, for
which $Z_\nu=\pm\sigma_\z$ in the case of SU(2), and we choose a list
of non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with
$\sum_\nu z_\nu=r$ to define the $r$-spin operator
\begin{align}
  X_z \equiv \bigotimes_{\nu=1}^n Z_\nu^{\otimes z_\nu}
  = \underbrace{Z_1 \otimes \cdots \otimes Z_1}_{z_1~\t{times}}
  \otimes
  \underbrace{Z_2 \otimes \cdots \otimes Z_2}_{z_2~\t{times}}
  \otimes \cdots \otimes
  \underbrace{Z_n \otimes \cdots \otimes Z_n}_{z_n~\t{times}}.
\end{align}
Our starting basis for $\M_r$ then consists of states
$\ket{m,\m w,X_z}\equiv\O\p{\m w,X_z}\ket{m}$, where we enforce that
$z_\nu\le m_\nu$ for all $\nu$ (where $m_\nu$ is the occupation number
of the single-spin state $\nu$), and that $z$ is zero in an entry
where $m$ is maximal: $z_\nu=0$ for
$\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  This second restriction is
equivalent to removing an operator from $\set{Z_\nu:\nu\in\ZZ_n}$,
making the remaining set linearly independent of the identity
operator.  The basis $\set{\ket{m,\m w,X_z}}$ for $\M_r$ is generally
complete, but not orthonormal, so as a final step we use the states
$\ket{m,\m w,X_z}$ to (numerically) construct an orthonormal basis
$\B_r$ for $\M_r$ (see Appendix \ref{sec:operator_content}).

@@

Next, we address the fact that the basis $\set{\ket{m,\m w,z}}$ for
$\M_r$ is overcomplete.  This fact is a simple consequence of our
choice of single-spin operators for the construction of
$\ket{m,\m w,z}$, namely $\set{Z_\nu:\nu\in\ZZ_n}$.  In essence, our
problem is the fact that $Z_\nu=\p{n-2}-\sum_{\mu\ne\nu}Z_\mu$, which
implies that any state $\ket{m,\m w,z}\in\M_r$ can be written as a
linear combination of states in which $z_\nu\to z_\nu-1$, and possibly
$z_\mu\to z_\mu+1$ for some $\mu\ne\nu$.  We can eliminate this
redundancy by removing an operator from the $\set{Z_\nu}$, or
equivalently enforcing that $z_\nu=0$ for some $\nu$.  As a matter of
choice, we therefore consider only states $\ket{m,\m w,z}$ for which
$z$ is zero at the first entry where $m$ is maximal: $z_\nu=0$ for
$\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  Given our previous restriction
that $z_\nu\le m_\nu$ for all $\nu$, enforcing that $z$ is zero at an
entry where $m$ is maximal maximizes the number of candidate values of
$z$ that we eliminate from consideration.

@@

To simplify $\Q\p{\m w,\m X}$, we first collect terms to express
\eqref{eq:sym_prod_start} as a sum of products:
\begin{align}
  \Q\p{\m w,\m X} = \sum_{k\in\D_N\p{\m M}} \prod_{\alpha\in\ZZ_p}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
  \label{eq:sym_prod_sum}
\end{align}
where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the dimensions
$M_\alpha$; and
$\D_N\p{\m M}\equiv\bigotimes_{M_\alpha\in\m M}\D_N\p{M_\alpha}$, such
that each $k\in\D_N\p{\m M}$ has the decomposition
$k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\D_N\p{M_\alpha}$.  We
then classify the terms in \eqref{eq:sym_prod_sum} by the numbers
$g_S$ of indices shared by all tensors $w_\alpha$ with
$\alpha\in S\subset\ZZ_p$.  For example, a term of the form
$w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
$a,b,c,d,e$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,3}} = 1,
  &&
  g_{\set{2,3}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be associated with the Venn diagram
\begin{align}
  \diagram{example_123},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
circles $\alpha\in S$.  We denote the set of all nonempty subsets of
$\ZZ_p$ by $\PP^*\p{\ZZ_p}$, and denote a choice of $g_S$ for all
$S\in\PP^*\p{\ZZ_p}$ by $g$.  With representations such as
\eqref{eq:venn_diagram} in mind, we will loosely refer to $g$ as
simply a ``diagram''.  We define
$\abs{g}\equiv\sum_{S\in\PP^*\p{\ZZ_p}}g_S$ to be the number of
indices considered in $g$ (or equivalently the number of dots in the
diagram $g$), and we denote the set of all diagrams for
$p=\dim\p{\m M}$ tensors with dimensions $\m M$ by $\G\p{\m M}$.  For
consistency, all diagrams $g\in\G\p{\m M}$ must associate exactly
$M_\alpha$ indices with the dimension-$M_\alpha$ tensor $w_\alpha$,
i.e.
\begin{align}
  \sum_{S\in\PP^*\p{\ZZ_p}\,:\,\alpha\in S} g_S
  = \sum_{R\in\PP^*\p{\ZZ_p\setminus\set{\alpha}}} g_{\set{\alpha}\cup R}
  = M_\alpha
\end{align}
for all $\alpha\in\ZZ_p$.  A classification of the terms in
\eqref{eq:sym_prod_sum} by diagrams $g\in\G\p{\m M}$ allows us to
expand
\begin{align}
  \Q\p{\m w,\m X} \EQPS \sum_{g\in\G\p{\m M}} \Q_g\p{\m w,\m X},
  \label{eq:sym_prod_diagrams}
\end{align}
where $\EQPS$ denotes equality under a restriction to $\M_0$, and
$\Q_g\p{\m w,\m X}$ is the sum of all terms in \eqref{eq:sym_prod_sum}
that are associated with the diagram $g$.
